Robustly decorrelating errors with mixed quantum gates

### 2019-08-21 decorrelation/mixed_unitaries.tex

\begin{equation}\label{eq:minimization}
  \begin{split}
    &\underset{\omega_i\geq0, |\omega|_1=1}{\textbf{minimize}: } ||\bold{L}\omega||_1\\
  \end{split}
\end{equation}
where $||\cdot||_1$ is the $\ell_1$ norm.

@@

% ==============================================================================
% Section: Robustness
% ==============================================================================
\subsection{Additional Constraints}
\subsubsection{Adding robustness} % (fold)
\label{sub:adding_robustness}
While mixed quantum channels offer significant improvements to gate performance, they fail to take into account the reality that most control electronics experience drift over time scales relevant to QIP performance. Because of this drift, the quality of the MQG will degrade. Thus, we would like to design MQGs that are robust to this drift. To enforce robustness, we can consider the generators in (\ref{eq:campbell-condition}) to be functions of a control vector $\vec{\delta}$, and instead of requiring that the error generators average to zero, we can impose a similar condition on the higher-order derivatives with respect to parameters that may drift. Specifically, we are interested in the $n^{th}$ order derivatives:
\begin{equation}
\label{eq:derivative}
D^n_i = \frac{1}{n!}\frac{\partial^{n}}{\partial\delta_{j_1}\ldots\partial\delta_{j_n}}\mathcal{L}_i(\vec{\delta})|_{\vec{\delta}=\vec{0}}
\end{equation}
If the dimension of $\vec{\delta}$ is $d$, the indices $j_0, \ldots, j_n$ take on values in $1, \ldots, d$ and this matrix has $d^n$ entries. 
We say that an MQG is robust to order $\ell$ (an $\ell$MQG) if for all $1 \leq k \leq \ell$:
\begin{equation}\label{eq:MQG}
\begin{gathered}
\sum_i\omega_i(\sum_{n=0}^k D^n_i)^n = 0\\
\end{gathered}
\end{equation}
In particular, we see that a 0MQG satisfies (\ref{eq:campbell-condition}). More generally, these conditions imply that an $\ell$MQG is insensitive to $\ell^{th}$ order in drift in $\vec{\delta}$. To see this, we can rewrite the implementations of the target gate as:
\begin{equation}\label{eq:taylor}
\begin{gathered}
\tilde{\mathcal{G}_i}(d\vec{\delta}) = \exp(-i(\mathcal{L}_i(\vec{0}) + \frac{\partial}{\partial\delta_j}\mathcal{L}_i(\vec{0})d\delta_j\\ +  \frac{1}{2}\frac{\partial^2}{\partial\delta_j\partial\delta_k} \mathcal{L}_i(\vec{0})d\delta_j d\delta_k + \ldots))\mathcal{G}
\end{gathered}
\end{equation}
By Taylor expanding (\ref{eq:taylor}) in $d\vec{\delta}$, one can see that the the first $\ell$ derivatives of an $\ell$MQG will be zero. Furthermore if we can find controls such that $||D_i^n||\leq\alpha$, (\ref{eq:MQG}) is approximately satisfied in the sense that:
\begin{equation}\label{eq:MQG-relaxed}
\begin{gathered}
||\sum\omega_iD^n_i||+ \mathcal{O}(\alpha^2) = 0\\
\end{gathered}
\end{equation}
where $||\cdot||$ is the Frobenius norm. This condition guarantees, as before, that errors will be suppressed to second order in $\alpha$, but now for all derivatives up to order $\ell$. %A proof is included in the appendix that generalizes the Hastings-Campbell Mixing Lemma in \cite{Campbell2017}. Namely, Campbell showed that if $0\in $ Conv$[\{G_i(\vec{\delta})\}]$, where Conv is the convex hull of its arguments, then an $\vec{\omega}$ exists that quadratically decreases the diamond norm. We prove that $0\in $ Conv$[\{D_j^n(\vec{\delta})\}]$ implies there is an $\vec{\omega}$ exists that quadratically decreases the $\ell^{th}$-order sensitivity of the diamond norm of an $\ell$MQG.
As discussed in \cite{Campbell2017}, the geometric picture for 0MQGs is to consider the error generators as vectors, and to ensure zero is contained in their convex hull. Figure \ref{fig:vectorspace} generalizes this picture to 1MQGs and gives intuition for the conditions required to produce them. To generate $\ell$MQGs, we first define the vectorized derivative matrix ${\bold{D}^{\ell}}$ in a similar way to (\ref{eq:vectorized_generators}):
\begin{equation}
{\bold{D}^{\ell}} =  \left(\begin{array}{cccc}
		D^\ell_{1_{1,1}} & D^\ell_{2_{1,1}} & \ldots   \\ 
		\vdots\ & \ddots &    \\
		D^\ell_{1_{n,n}} &  &  D^\ell_{m_{n,n}} \\ 
	\end{array} 	
	\right)
\end{equation}

@@

\subsubsection{Hamiltonian Norm Regularization}
\label{sec:norm}
The convex optimization problems discussed up until this point can be solved analytically, however casting them as convex optimization problems will allow us to add penalty terms to the cost function to promote different properties of the MQG. In particular, while (\ref{eq:robust_minimization}) is sufficient for suppressing the diamond norm to first order relative to the \textit{worst} controls in the collection, it does not preferentially select the controls with the least error. As an example, any MQG that mixes equally over equal and opposite errors will satisfy (\ref{eq:robust_minimization}), but the magnitude of the error of the resulting channel will depend on the magnitude of the original coherent errors. To encourage the inclusion of controls with smaller error, we may impose a penalty proportional to the norm of the included Hamiltonians. In our case, we choose to penalize for the $\ell_2$ ($||\cdot||_2$) norm:

@@

In many machine learning contexts, lasso regularization \cite{tibshirani1996regression} can be used to enforce sparsity in solutions, however this is insufficient for our purposes, as we already constrain $\omega$ to be a valid probability distribution. The problem of enforcing sparsity in such situations has been considered in \cite{NIPS2012_4504} and can be expressed via another convex program that extends (\ref{eq:robust_minimization}):

@@

\subsection{Experimental} % (fold)
\label{sub:experimental}
Here we present experimental results from implementing this routine on a fixed-frequency, superconducting transmon qubit. In particular, we use qubit 8 on the Rigetti 19Q-Acorn chip, whose characterization can be found in \cite{1712.05771}. To implement an MQG on this qubit, four miscalibrated Gaussian pulses were produced by scaling the amplitude for a calibrated 10 sample 50ns $RX(\frac{\pi}{2})$ pulse by $106.4\%$,  $103.9\%$, $93.7\%$ and $91.2\%$.

@@

More importantly, minimizing the off-diagonal elements of the process matrix, we expect to produce a process with suppressed coherent error. To see that this is the case, we can observe the drastic reduction in the variance at each point for the MQG relative to the miscalibrated pulses. A discussed in \cite{Ball2016}, for non-Markovian error models, noise will manifest as gamma distributed points for each sequence length. On the other hand, Markovian noise, such as depolarizing noise, will result in Gaussian distributed fidelity estimates for each randomized benchmarking sequence length. We see that the coherently miscalibrated controls in our RB experiment have long tails, consistent with gamma distributed random variables, while the calibrated and randomized implementations both have much shorter tails, consistent with Gaussian distributed random variables. Thus our experiment demonstrates that the MQG has a significantly less-coherent error channel.
% subsection experimental (end)

% ==============================================================================
% Section: Numerical Results
% ==============================================================================
\subsection{Numerical Implementation}
\label{sec:numerical_results}
In the following numerical results, we use the methods in Section \ref{sec:mixed_unitary_processes} to build MQGs. We consider the following model for a single tunable qubit: 
\begin{equation}\label{eq:1Qham}
  H(\delta, \epsilon, t) = \epsilon\sigma_z + (1 + \delta)(c_x(t)\sigma_x + c_y(t)\sigma_y)
\end{equation}
where $\epsilon$ is a random variable used to model fluctuations in the qubit frequency, and $\delta$ is a random variable used to model fluctuations in the control field. We use the GRAPE algorithm\cite{Khaneja2005} with N=25 steps and total evolution time of $\pi$ to generate 100 candidate controls.

@@

Using controls generated in this way, the MQGs produced for $RX(\frac{\pi}{2})$ and $RY(\frac{\pi}{2})$ are qualitatively similar, with the results for $RY(\frac{\pi}{2})$ shown in Figure \ref{fig:YMQG}. By imposing the $\ell_2$ penalty from Section \ref{sec:norm}, so that the algorithm preferentially selects controls with smaller errors. Adding this constraint made the 0MQG perform better at the origin by nearly an order of magnitude, and moved the 1MQG from being out-performed by the 0MQG for all detunings, to out performing by nearly an order of magnitude when there is $.1\%$ drift in the controls. Imposing this constraint allows us to trade off flatness at the origin for performance at the origin. This shows that through adding constraints to our optimization routine, we can make the MQG practically useful.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{SQRTY_no_member.png}
  \caption{Numerical results comparing a 0MQG to a 1MQG for a single tunable qubit, for $RY(\frac{\pi}{2})$. The results are qualitatively similar to those for $RX(\frac{\pi}{2})$. In this case the 0MQG outperforms both the 1MQG by two orders of magnitude, and the constituent controls by three orders of magnitude at the origin. However, varying over $\delta$ we see that the 1MQG outperforms the 0MQG by up to an order of magnitude when there is $.1\%$ drift in the qubit control amplitudes.}
  \label{fig:YMQG}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{2QRBC_no_member.png}
  \caption{Numerical results comparing a 0MQG to a 1MQG for a pair of tunable qubits, with a resonant exchange interaction. Shown with lower alpha values are example constituent controls. The 0MQG and 1MQG can be seen to outperform these controls by half of an order of magnitude at the origin. For all detuning values the 1MQG performs as well or better than the 0MQG. When there is $.2\%$ drift in the qubit frequency, the 1MQG outperforms members of the control families by almost an order of magnitude in diamond norm. Similarly, for $.2\%$ drift in the qubit control amplitude, we see that the 1MQG outperforms the the constituent controls by over half an order of magnitude.}
  \label{fig:2MQG}
\end{figure}

>>>>

\begin{equation}\label{eq:minimization}
  \begin{split}
    &\underset{\omega_i\geq0, |\omega|_1=1}{\textbf{minimize}: } ||\bold{L}\omega||_2\\
  \end{split}
\end{equation}
where $||\cdot||_2$ is the $\ell_2$ norm.

@@

% ==============================================================================
% Section: Robustness
% ==============================================================================
\subsection{Additional Constraints}
\subsubsection{Adding robustness} % (fold)
\label{sub:adding_robustness}
While mixed quantum channels offer significant improvements to gate performance, they fail to take into account the reality that most control electronics experience drift over time scales relevant to QIP performance. Because of this drift, the quality of the MQG will degrade. Thus, we would like to design MQGs that are robust to this drift. To enforce robustness, we can consider the generators in (\ref{eq:campbell-condition}) to be functions of a control vector $\vec{\delta}$, and instead of requiring that the error generators average to zero, we can impose a similar condition on the higher-order derivatives with respect to parameters that may drift. Specifically, we are interested in the $n^{th}$ order derivatives:
\kcy{move this later}
\begin{equation}
\label{eq:derivative}
D^n_i = \frac{1}{n!}\frac{\partial^{n}}{\partial\delta_{j_1}\ldots\partial\delta_{j_n}}\mathcal{L}_i(\vec{\delta})|_{\vec{\delta}=\vec{0}}
\end{equation}
If the dimension of $\vec{\delta}$ is $d$, the indices $j_0, \ldots, j_n$ take on values in $1, \ldots, d$ and this matrix has $d^n$ entries. 
We say that an MQG is robust to order $\ell$ (an $\ell$MQG) if for all $1 \leq k \leq \ell$:
\begin{equation}\label{eq:MQG}
\begin{gathered}
\sum_i\omega_i(\sum_{n=0}^k D^n_i)^n = 0\\
\end{gathered}
\end{equation}
In particular, we see that a 0MQG satisfies (\ref{eq:campbell-condition}). More generally, these conditions imply that an $\ell$MQG is insensitive to $\ell^{th}$ order in drift in $\vec{\delta}$. To see this, we can rewrite the implementations of the target gate as:
\kcy{move this earlier}
\begin{equation}\label{eq:taylor}
\begin{gathered}
\tilde{\mathcal{G}_i}(d\vec{\delta}) = \exp(-i(\mathcal{L}_i(\vec{0}) + \frac{\partial}{\partial\delta_j}\mathcal{L}_i(\vec{0})d\delta_j\\ +  \frac{1}{2}\frac{\partial^2}{\partial\delta_j\partial\delta_k} \mathcal{L}_i(\vec{0})d\delta_j d\delta_k + \ldots))\mathcal{G}
\end{gathered}
\end{equation}
By Taylor expanding (\ref{eq:taylor}) in $d\vec{\delta}$, one can see that the the first $\ell$ derivatives of an $\ell$MQG will be zero. Furthermore if we can find controls such that $||D_i^n||\leq\alpha$, (\ref{eq:MQG}) is approximately satisfied in the sense that:
\begin{equation}\label{eq:MQG-relaxed}
\begin{gathered}
||\sum\omega_iD^n_i||+ \mathcal{O}(\alpha^2) = 0\\
\end{gathered}
\end{equation}
where $||\cdot||$ is the Frobenius norm. This condition guarantees, as before, that errors will be suppressed to second order in $\alpha$, but now for all derivatives up to order $\ell$. %A proof is included in the appendix that generalizes the Hastings-Campbell Mixing Lemma in \cite{Campbell2017}. Namely, Campbell showed that if $0\in $ Conv$[\{G_i(\vec{\delta})\}]$, where Conv is the convex hull of its arguments, then an $\vec{\omega}$ exists that quadratically decreases the diamond norm. We prove that $0\in $ Conv$[\{D_j^n(\vec{\delta})\}]$ implies there is an $\vec{\omega}$ exists that quadratically decreases the $\ell^{th}$-order sensitivity of the diamond norm of an $\ell$MQG.
As discussed in \cite{Campbell2017}, the geometric picture for 0MQGs is to consider the error generators as vectors, and to ensure zero is contained in their convex hull. Figure \ref{fig:vectorspace} generalizes this picture to 1MQGs and gives intuition for the conditions required to produce them. To generate $\ell$MQGs, we first define the vectorized derivative matrix ${\bold{D}^{\ell}}$ in a similar way to (\ref{eq:vectorized_generators}):
\begin{equation}
{\bold{D}^{\ell}} =  \left(\begin{array}{cccc}
		D^\ell_{1_{1,1}} & D^\ell_{2_{1,1}} & \ldots   \\ 
		\vdots\ & \ddots &    \\
		D^\ell_{1_{n,n}} &  &  D^\ell_{m_{n,n}} \\ 
	\end{array} 	
	\right)
\end{equation}

@@

\subsubsection{Hamiltonian Norm Regularization}
\label{sec:norm}
The convex optimization problems discussed up until this point can be solved analytically, \kcy{nope!} however casting them as convex optimization problems will allow us to add penalty terms to the cost function to promote different properties of the MQG. In particular, while (\ref{eq:robust_minimization}) is sufficient for suppressing the diamond norm to first order relative to the \textit{worst} controls in the collection, it does not preferentially select the controls with the least error. As an example, any MQG that mixes equally over equal and opposite errors will satisfy (\ref{eq:robust_minimization}), but the magnitude of the error of the resulting channel will depend on the magnitude of the original coherent errors. To encourage the inclusion of controls with smaller error, we may impose a penalty proportional to the norm of the included Hamiltonians. In our case, we choose to penalize for the $\ell_2$ ($||\cdot||_2$) norm:

@@

In many machine learning contexts, lasso regularization \cite{tibshirani1996regression} can be used to enforce sparsity in solutions, however this is insufficient for our purposes, as we already constrain $\omega$ to be a valid probability distribution. \kcy{more here} The problem of enforcing sparsity in such situations has been considered in \cite{NIPS2012_4504} and can be expressed via another convex program that extends (\ref{eq:robust_minimization}):

@@

\subsection{Experimental} % (fold)
\label{sub:experimental}
\kcy{more introduction} 
Here we present experimental results from implementing this routine on a fixed-frequency, superconducting transmon qubit. In particular, we use qubit 8 on the Rigetti 19Q-Acorn chip, whose characterization can be found in \cite{1712.05771}. To implement an MQG on this qubit, four intentionally miscalibrated Gaussian pulses were produced by scaling the amplitude of a calibrated, 10-sample, 50ns $RX(\frac{\pi}{2})$ pulse by $106.4\%$,  $103.9\%$, $93.7\%$ and $91.2\%$.

@@

\kcy{Kill all non-markovians}
More importantly, minimizing the off-diagonal elements of the process matrix, we expect to produce a process with suppressed coherent error. To see that this is the case, we can observe the drastic reduction in the variance at each point for the MQG relative to the miscalibrated pulses. A discussed in \cite{Ball2016}, for non-Markovian error models, noise will manifest as gamma distributed points for each sequence length. On the other hand, Markovian noise, such as depolarizing noise, will result in Gaussian distributed fidelity estimates for each randomized benchmarking sequence length. We see that the coherently miscalibrated controls in our RB experiment have long tails, consistent with gamma distributed random variables, while the calibrated and randomized implementations both have much shorter tails, consistent with Gaussian distributed random variables. Thus our experiment demonstrates that the MQG has a significantly less-coherent error channel.
% subsection experimental (end)

% ==============================================================================
% Section: Numerical Results
% ==============================================================================
\subsection{Numerical Implementation}
\label{sec:numerical_results}
\kcy{more introduction - what is going on. We are constructing solutions that are themseleves naturally robust, then combining them to make a MQG that is... something better? }
In the following numerical results, we use the methods in Section \ref{sec:mixed_unitary_processes} to build MQGs. We consider the following model for a single tunable qubit: 
\begin{equation}\label{eq:1Qham}
  H(\delta, \epsilon, t) = \epsilon\sigma_z + (1 + \delta)(c_x(t)\sigma_x + c_y(t)\sigma_y)
\end{equation}
\kcy{dimensionless}
where $\epsilon$ is a random variable used to model fluctuations in the qubit frequency, and $\delta$ is a random variable used to model fluctuations in the control field. We use the GRAPE algorithm\cite{Khaneja2005} with N=25 steps and total evolution time of $\pi$ to generate 100 candidate controls.

@@

Using controls generated in this way, the MQGs produced for $RX(\frac{\pi}{2})$ and $RY(\frac{\pi}{2})$ are qualitatively similar, with the results for $RY(\frac{\pi}{2})$ shown in Figure \ref{fig:YMQG}. \kcy{not a sentence -->} By imposing the $\ell_2$ penalty from Section \ref{sec:norm}, so that the algorithm preferentially selects controls with smaller errors. Adding this constraint made the 0MQG perform better at the origin by nearly an order of magnitude, and moved the 1MQG from being out-performed by the 0MQG for all detunings, to out performing by nearly an order of magnitude when there is $.1\%$ drift in the controls. Imposing this constraint allows us to trade off flatness at the origin for performance at the origin. This shows that through adding constraints to our optimization routine, we can make the MQG practically useful.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{SQRTY_no_member.png}
  \caption{\kcy{Axis label is screwed up} Numerical results comparing a 0MQG to a 1MQG for a single tunable qubit, for $RY(\frac{\pi}{2})$. The results are qualitatively similar to those for $RX(\frac{\pi}{2})$. In this case the 0MQG outperforms both the 1MQG by two orders of magnitude, and the constituent controls by three orders of magnitude at the origin. However, varying over $\delta$ we see that the 1MQG outperforms the 0MQG by up to an order of magnitude when there is $.1\%$ drift in the qubit control amplitudes.}
  \label{fig:YMQG}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{2QRBC_no_member.png}
  \caption{\kcy{needs to be on previous page} Numerical results comparing a 0MQG to a 1MQG for a pair of tunable qubits, with a resonant exchange interaction. Shown with lower alpha values are example constituent controls. The 0MQG and 1MQG can be seen to outperform these controls by half of an order of magnitude at the origin. For all detuning values the 1MQG performs as well or better than the 0MQG. When there is $.2\%$ drift in the qubit frequency, the 1MQG outperforms members of the control families by almost an order of magnitude in diamond norm. Similarly, for $.2\%$ drift in the qubit control amplitude, we see that the 1MQG outperforms the the constituent controls by over half an order of magnitude.}
  \label{fig:2MQG}
\end{figure}
